## Reference Hydrometric Basin Network

The data tidying, calculation of Indicators of Hydrologic Alteration (IHA) variables, ice variables, and freshet initiation for ~1000 reference stations in watersheds across Canada in order to assess Cold-Water Hydrologic Indicators of Change (CHIC) variables.

Jacqui Levy

```{r, echo=FALSE, include = FALSE, warning= FALSE}
library(IHA)
library(tidyverse)
library(tidyhydat)
library(zoo)
library(lubridate)
library(ggplot2)
library(dataRetrieval)
library(Kendall)
library(knitr)
library(plotly)
library(data.table)
library(manipulateWidget)

#install.packages("IHA", repos="http://R-Forge.R-project.org")

```


##RHBN stations 
```{r}

stns <- read.csv("data/RHBN_Metadata.csv" ) %>% filter(DATA_TYPE == "Q") 

```


```{r}

#sanity checks
n_distinct(stns$STATION_NUMBER) #1058
n_distinct(allstns$STATION_NUMBER) #1057

stns_ids <- stns %>% select(STATION_NUMBER) %>% distinct                                
allstns_ids <- allstns %>% select(STATION_NUMBER) %>% distinct

anti_join(stns_ids, allstns_ids)
# station 08LD001 not in master list of all stations, need to get this one or find out if need it or not 

```


```{r}

#getting station numbers by ecoregion into a list 

stn_num_eco <- stns %>% 
  mutate(Ecoregion = ifelse(Ecoregion == '', "No Ecoregion", Ecoregion)) %>%
  distinct(STATION_NUMBER, .keep_all = TRUE) %>%
  group_by(Ecoregion) %>%
  summarise(STATION_NUMBER = list(STATION_NUMBER))

#loop through this and get several vectors of station ids

ecoreglist <- list()
for (i in unique(stn_num_eco$Ecoregion)) {
    
    df_subset <- stn_num_eco[stn_num_eco$Ecoregion == i,]
    vec <-  as.vector(unlist(df_subset$STATION_NUMBER))
    names(vec) <- i
    ecoreglist[[i]] <- vec
}

#get separate dfs for flow for each ecoregion
#takes too much memory, don't use for now - consider google collab?
 # for (i in names(ecoreglist)){
#  stn_num <- unlist(ecoreglist$i)
#  df <- tidyhydat::hy_daily_flows(station_number = stn_num)
 # assign(paste('',i,sep=''), df)
  #names(ecoreglist) <- i
  }

```


```{r}

#Get dfs of daily flows for all ecoregions
eastern_temp_forests <- tidyhydat::hy_daily_flows(station_number = ecoreglist$`Eastern Temperate Forests`)
great_plains <- tidyhydat::hy_daily_flows(station_number = ecoreglist$`Great Plains`)
hudson_plains <- tidyhydat::hy_daily_flows(station_number = ecoreglist$`Hudson Plains`)
w_coast_forests <- tidyhydat::hy_daily_flows(station_number = ecoreglist$`Marine West Coast Forests`)
no_ecoregion <- tidyhydat::hy_daily_flows(station_number = ecoreglist$`No Ecoregion`)
n_american_deserts <- tidyhydat::hy_daily_flows(station_number = ecoreglist$`North American Deserts`)
northern_forests <- tidyhydat::hy_daily_flows(station_number = ecoreglist$`Northern Forests`)
nw_forested_mtns <- tidyhydat::hy_daily_flows(station_number = ecoreglist$`Northwest Forested Mountains`)
taiga <- tidyhydat::hy_daily_flows(station_number = ecoreglist$Taiga)
tundra <- tidyhydat::hy_daily_flows(station_number = ecoreglist$Tundra)

rm(ecoreglist, stn_num_eco, df_subset)

```

```{r}

#put all dfs in a list

df_list <- list(eastern_temp_forests, great_plains, hudson_plains, w_coast_forests, no_ecoregion, n_american_deserts, northern_forests, nw_forested_mtns, taiga, tundra)
names(df_list) <- c('eastern_temp_forests', 'great_plains', 'hudson_plains', 'w_coast_forests', 'no_ecoregion', 'n_american_deserts', 'northern_forests', 'nw_forested_mtns', 'taiga', 'tundra')

```

#tidying methods and codes adapted from eflows project

```{r}
source("scripts/Eflows_FUNCTIONS.R")

#create a smaller practice/test df for developing workflows
df_lst <- list(taiga, tundra)
names(df_lst) <- c("taiga", "tundra")

#note: need to source all the functions in another script

#tidying function
tidying <- function(df){
  tidied <- df %>%
  group_by(STATION_NUMBER) %>%
  tidyr::complete(Date = seq.Date(as.Date("1969/10/1"), as.Date("2022/09/30"), by="day")) %>%
  mutate(Date = as.Date(Date)) %>%
  filter(Date >= "2018-10-01" & Date < "2023-10-01") #change this later on
  return(tidied)
}

df_lst <- map(df_lst, tidying)


#get df of distinct stations with desired columns
stns_cs <- stns %>% 
  mutate(Ecoregion = ifelse(Ecoregion == '', "No Ecoregion", Ecoregion)) %>%
  distinct(STATION_NUMBER, .keep_all = TRUE) %>%
  select(STATION_NUMBER, RHBN.N, Ecoregion, Data.Quality, Years.C, Years.S)

#add the seasonal vs continuous data to the dfs in the list of dfs 
for (i in seq_along(df_lst)){
  df_lst[[i]] <- inner_join(df_lst[[i]], stns_cs, by = "STATION_NUMBER")
}

#add water year to all stations 
df_lst <- map(df_lst, addWaterYear)


#make a loop to add a day of the year and return a list of dfs, separated by ecoregion
#(function doesn't work with map, need to open issue and fix for icicle r package)
df_list <- list()
for (i in seq_along(df_lst)) {
  df_subset <- df_lst[[i]] #subset of df, by ecoregion
  stn_all_split <- split(df_subset, df_subset$STATION_NUMBER) #split by station number to group them together
  stn_all_split_doy <- lapply(stn_all_split, calc_day_of_wyear) #apply day of year function to each station
  stn_all <- bind_rows(stn_all_split_doy, .id = "STATION_NUMBER") #bind the stations back into one dataframe
  ecoregion <- stn_all$Ecoregion[i] #find the name of the ecoregion for that group of stations
  df_list[[i]] <- stn_all     #append the new df to a list
  names(df_list)[i] <- c(ecoregion) #name the list of dfs by their corresponding ecoregion
}

```


```{r}

#create a function to filter out the seasonal months for seasonal stations only (Data.Quality =)
filter_seas <- function(data) {
  df <- {{data}} %>%
  filter(Data.Quality == 'S' & month(Date) %in% c(3,4,5,6,7,8,9,10))
}

df_lst <- map(df_lst, filter_seas)

```


#Plotting part 1
```{r}

#make a bunch of dfs for plotting and then put them in a list 
taiga_test_yrs <- df_lst$Taiga %>%
  dplyr::group_by(STATION_NUMBER) %>%
  dplyr::summarise(Year = unique(waterYear)) %>%
  arrange(STATION_NUMBER)

tundra_test_yrs

c <- ceiling(n_distinct(unique(df_lst$Taiga$STATION_NUMBER))/30)

df_parts <- split(taiga_test_yrs, cut(seq_len(nrow(taiga_test_yrs)), breaks = c))

taiga1 <- df_parts[[1]]
taiga2 <- df_parts[[2]]
taiga3 <- df_parts[[3]]

df_list_toplot <- list(taiga1, taiga2, taiga3, tundra_test_yrs)
names(df_list_toplot) <- c('taiga1', 'taiga2', 'taiga3', 'tundra_test_yrs')

```

Plotting part 2
```{r}
#plot data availability from list of dfs above

for (i in seq_along(df_list_toplot)) {
  df_subset <- df_list_toplot[[i]] #subset of df, by ecoregion

#ecoregion <- df_list_toplot[i] #find the name of the ecoregion for that group of stations

#yrs_full <- df_subset %>%
  #dplyr::group_by(STATION_NUMBER) %>%
  #dplyr::summarise(Year = unique(waterYear))

p<- ggplot(df_subset, aes(x = Year, y = STATION_NUMBER)) +
  geom_tile(fill = "blue", color = "white") +
  theme_classic() +
  ylab("") +
 ggtitle(paste("Data Availability - ", names(df_list_toplot))[i]) +
    theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1, size= 9), plot.title = element_text(hjust = 0.5)) 
print(p)
}

```



Delete years with >3 missing days

```{r}

#delete missing data
df_lst_del_yrs <-map(df_lst, calc_rle)


```

#Plotting post-deleting years
```{r}
#make a bunch of dfs for plotting and then put them in a list 
taiga_test_del_yrs <- df_lst_del_yrs$Taiga %>%
  dplyr::group_by(STATION_NUMBER) %>%
  dplyr::summarise(Year = unique(waterYear)) %>%
  arrange(STATION_NUMBER)

tundra_test_del_yrs <- df_lst_del_yrs$Tundra %>%
  dplyr::group_by(STATION_NUMBER) %>%
  dplyr::summarise(Year = unique(waterYear)) %>%
  arrange(STATION_NUMBER)


c <- ceiling(n_distinct(unique(df_lst$Taiga$STATION_NUMBER))/30)

df_parts <- split(taiga_test_yrs, cut(seq_len(nrow(taiga_test_yrs)), breaks = c))

taiga1 <- df_parts[[1]]
taiga2 <- df_parts[[2]]
taiga3 <- df_parts[[3]]

df_list_toplot_del <- list(taiga1, taiga2, taiga3, tundra_test_yrs)
names(df_list_toplot_del) <- c('taiga1', 'taiga2', 'taiga3', 'tundra_test_del_yrs')



```

#plotting post deleting years
```{r}

#plot data availability from list of dfs above

for (i in seq_along(df_list_toplot_del)) {
  df_subset <- df_list_toplot_del[[i]] #subset of df, by ecoregion

#ecoregion <- df_list_toplot[i] #find the name of the ecoregion for that group of stations

#yrs_full <- df_subset %>%
  #dplyr::group_by(STATION_NUMBER) %>%
  #dplyr::summarise(Year = unique(waterYear))

p<- ggplot(df_subset, aes(x = Year, y = STATION_NUMBER)) +
  geom_tile(fill = "orange", color = "white") +
  theme_classic() +
  ylab("") +
 ggtitle(paste("Data Availability - ", names(df_list_toplot_del))[i]) +
    theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1, size= 9), plot.title = element_text(hjust = 0.5)) 
print(p)
}


```


```{r}


```


```{r}


```

#IHA
```{r}

#tidying to do before IHA:
mutate(waterYear = as.character(waterYear))
  mutate(Date = as.Date(Date)) %>%
  filter(waterYear > 1970) %>%
  filter(waterYear < 2023)

#reformat date and fill remaining missing values with preceding values
stns_daymonthyear <- stns_ready %>%
  mutate(Date = format(Date,"%d-%m-%Y")) %>%
  group_by(STATION_NUMBER) %>%
  fill(Value, .direction = 'downup') #maybe don't do this in this case?

```
 
```{r}



```

```{r}



```

```{r}

#next find all years available for each station
allstns <- df_list %>%
mutate(Date = ymd(Date),  Year = format_ISO8601(Date, precision = "y"))
filter(Year > 1969 & Year < 2024)

allstns_dt %>%
  group_by(STATION_NUMBER) %>%
  summarize(years_with_data =toString(unique(Year)))

#years with good data - delete years with >3 days a month of missing data, delete that whole year before doing calculations 


```

```{r}
#graph the data availability for all of the stations before deleting years with 3 days of data missing

graph_func <- function(data){
  ecoregion <- {{data}}[i]

yrs_full <- {{data}} %>%
  dplyr::group_by(STATION_NUMBER) %>%
  dplyr::summarise(Year = unique(waterYear))


#graphing data availability
p <- ggplot(yrs_full, aes(x = Year, y = STATION_NUMBER)) +
  geom_tile(fill = "lightgreen", color = "white") +
  theme_classic() +
  ylab("") +
  ggtitle("Data Availability - ") +
    theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1, size= 9), plot.title = element_text(hjust = 0.5))
return(p)
}


map(df_lst, graph_func) #cant title by the name of each df if do it this way, need to use a loop



```

```{r}


ggplot(diamonds, aes(carat, price)) + 
  geom_hex()
ggsave("output/diamonds.pdf")

write_csv(diamonds, "output/diamonds.csv")

summarise(STATION_NUMBER = paste(STATION_NUMBER,collapse = "','"))


```

