## Reference Hydrometric Basin Network

The data tidying, calculation of Indicators of Hydrologic Alteration (IHA) variables, ice variables, and freshet initiation for ~1300 reference stations in watersheds across Canada in order to assess Cold-Water Hydrologic Indicators of Change (CHIC) variables.

Jacqui Levy

```{r, echo=FALSE, include = FALSE, warning= FALSE}

library(IHA)
library(tidyverse)
library(tidyhydat)
library(zoo)
library(lubridate)
library(ggplot2)
library(dataRetrieval)
library(Kendall)
library(knitr)
library(plotly)
library(data.table)
library(manipulateWidget)

#install.packages("IHA", repos="http://R-Forge.R-project.org")

```


##Get all RHBN stations 
```{r}
#from govt canada RHBN list: https://www.canada.ca/en/environment-climate-change/services/water-overview/quantity/monitoring/survey/data-products-services/reference-hydrometric-basin-network.html

stns <- read.csv("data/RHBN_Metadata.csv" ) %>% filter(DATA_TYPE == "Q") 

```


```{r}

#put station numbers into a list, by ecoregion

stn_num_eco <- stns %>% 
  mutate(Ecoregion = ifelse(Ecoregion == '', "No Ecoregion", Ecoregion)) %>%
  distinct(STATION_NUMBER, .keep_all = TRUE) %>%
  group_by(Ecoregion) %>%
  summarise(STATION_NUMBER = list(STATION_NUMBER))

#loop through this and get several vectors of station ids

ecoreglist <- list()
for (i in unique(stn_num_eco$Ecoregion)) {
    
    df_subset <- stn_num_eco[stn_num_eco$Ecoregion == i,]
    vec <-  as.vector(unlist(df_subset$STATION_NUMBER))
    names(vec) <- i
    ecoreglist[[i]] <- vec
}


```


```{r}

#Use tidyhydat database to get daily flows for all ecoregions and put all dataframes in a list
eastern_temp_forests <- tidyhydat::hy_daily_flows(station_number = ecoreglist$`Eastern Temperate Forests`)
great_plains <- tidyhydat::hy_daily_flows(station_number = ecoreglist$`Great Plains`)
hudson_plains <- tidyhydat::hy_daily_flows(station_number = ecoreglist$`Hudson Plains`)
w_coast_forests <- tidyhydat::hy_daily_flows(station_number = ecoreglist$`Marine West Coast Forests`)
no_ecoregion <- tidyhydat::hy_daily_flows(station_number = ecoreglist$`No Ecoregion`)
n_american_deserts <- tidyhydat::hy_daily_flows(station_number = ecoreglist$`North American Deserts`)
northern_forests <- tidyhydat::hy_daily_flows(station_number = ecoreglist$`Northern Forests`)
nw_forested_mtns <- tidyhydat::hy_daily_flows(station_number = ecoreglist$`Northwest Forested Mountains`)
taiga <- tidyhydat::hy_daily_flows(station_number = ecoreglist$Taiga)
tundra <- tidyhydat::hy_daily_flows(station_number = ecoreglist$Tundra)



#put all dfs in a list

df_list <- list(eastern_temp_forests, great_plains, hudson_plains, w_coast_forests, no_ecoregion, n_american_deserts, northern_forests, nw_forested_mtns, taiga, tundra)
names(df_list) <- c('eastern_temp_forests', 'great_plains', 'hudson_plains', 'w_coast_forests', 'no_ecoregion', 'n_american_deserts', 'northern_forests', 'nw_forested_mtns', 'taiga', 'tundra')

rm(ecoreglist, stn_num_eco, df_subset)

rm(eastern_temp_forests, great_plains, hudson_plains, w_coast_forests, no_ecoregion, n_american_deserts, northern_forests, nw_forested_mtns, taiga, tundra )

```


## Tidying
methods and codes adapted from eflows project

```{r}
source("scripts/Eflows_FUNCTIONS.R")

#note: need to source the rest of the tidying functions in a .R script

#tidying function - sequence full set of years
tidying <- function(df){
  tidied <- df %>%
  group_by(STATION_NUMBER) %>%
  tidyr::complete(Date = seq.Date(as.Date("1969/10/1"), as.Date("2022/09/30"), by="day")) %>%
  mutate(Date = as.Date(Date)) %>%
  filter(Date >= "1970-10-01" & Date < "2023-10-01") 
  return(tidied)
}

df_list <- map(df_list, tidying)

#get df of distinct stations with desired columns
#probably don't need anymore 
stns_cs <- stns %>% 
  mutate(Ecoregion = ifelse(Ecoregion == '', "No Ecoregion", Ecoregion)) %>%
  distinct(STATION_NUMBER, .keep_all = TRUE) %>%
  select(STATION_NUMBER, RHBN.N, Ecoregion, Data.Quality, Years.C, Years.S)

#add the seasonal vs continuous data to the dfs in the list of dfs 
for (i in seq_along(df_list)){
  df_list[[i]] <- inner_join(df_list[[i]], stns_cs, by = "STATION_NUMBER")
}

#add water year to all stations 
df_list <- map(df_list, addWaterYear)

```

#Day of the Year
```{r}
#make a loop to add a day of the year and return a list of dfs, separated by ecoregion
#(doy function doesn't work with map, need to open issue and fix for icicle r package)

dflst <- list()
for (i in seq_along(df_list)) {
  df_subset <- df_list[[i]] #subset of df, by ecoregion
  stn_all_split <- split(df_subset, df_subset$STATION_NUMBER) #split by station number to group them together
  stn_all_split_doy <- lapply(stn_all_split, calc_day_of_wyear) #apply day of year function to each station
  stn_all <- bind_rows(stn_all_split_doy, .id = "STATION_NUMBER") #bind the stations back into one dataframe
  ecoregion <- stn_all$Ecoregion[i] #find the name of the ecoregion for that group of stations
  dflst[[i]] <- stn_all     #append the new df to a list
  names(dflst)[i] <- c(ecoregion) #name the list of dfs by their corresponding ecoregion
}

```

#Missing data 
zero flow days and NAs: by month and annually 

```{r}

missing_days <- function(data) {
  input_df <- {{data}}
   results_df <- input_df %>%
  dplyr::group_by(STATION_NUMBER, waterYear) %>%
  dplyr::summarize(Jan_missingdays = sum(is.na(Value) & month(Date) == 1),
            Feb_missingdays = sum(is.na(Value) & month(Date) == 2),
            Mar_missingdays = sum(is.na(Value) & month(Date) == 3),
            Apr_missingdays = sum(is.na(Value) & month(Date) == 4),
            May_missingdays = sum(is.na(Value) & month(Date) == 5),
            June_missingdays = sum(is.na(Value) & month(Date) == 6),
            July_missingdays = sum(is.na(Value) & month(Date) == 7),
            Aug_missingdays = sum(is.na(Value) & month(Date) == 8),
            Sept_missingdays = sum(is.na(Value) & month(Date) == 9),
            Oct_missingdays = sum(is.na(Value) & month(Date) == 10),
            Nov_missingdays = sum(is.na(Value) & month(Date) == 11),
            Dec_missingdays = sum(is.na(Value) & month(Date) == 12),
            Annual_missingdays = sum(is.na(Value)),
            Jan_zero_flow_days = sum(Value == 0 & month(Date) == 1),
            Feb_zero_flow_days = sum(Value == 0 & month(Date) == 2),
            Mar_zero_flow_days = sum(Value == 0 & month(Date) == 3),
            Apr_zero_flow_days = sum(Value == 0 & month(Date) == 4),
            May_zero_flow_days = sum(Value == 0 & month(Date) == 5),
            Jun_zero_flow_days = sum(Value == 0 & month(Date) == 6),
            Jul_zero_flow_days = sum(Value == 0 & month(Date) == 7),
            Aug_zero_flow_days = sum(Value == 0 & month(Date) == 8),
            Sep_zero_flow_days = sum(Value == 0 & month(Date) == 9),
            Oct_zero_flow_days = sum(Value == 0 & month(Date) == 10),
            Nov_zero_flow_days = sum(Value == 0 & month(Date) == 11),
            Dec_zero_flow_days = sum(Value == 0 & month(Date) == 12), 
            Annual_zero_flow_days = sum(Value == 0))       
   
   }

df_missing_yrs_months  <- map(dflst, missing_days)


```

#Delete missing data 
(by month)

```{r}

#delete any month with more than 3 missing days

remove_months <- function(df) {
  del_mnths <- df %>%
    group_by(STATION_NUMBER, waterYear, month=month(Date)) %>%
    filter(sum(is.na(Value)) <= 4) %>%
  ungroup() %>%
  select(-month)
return(del_mnths)
}

df_list_rem_months  <- map(dflst, remove_months)

```


#Anuual mean & median

```{r}

#Note: IHA group 1 currently measures the monthly median. Can switch to FUN = mean to get monthly means. 

calc_mn_md <- function(data){
  input_data <- {{data}}
  results <- input_data %>%
    dplyr::group_by(STATION_NUMBER, waterYear) %>%
    dplyr::summarize(annualmean = mean(Value), annualmedian = median(Value))
}

df_yrly_means_med  <- map(df_list_rem_months, calc_mn_md)


```


#IHA

```{r}


tidying_iha <- function(df){
  #mutate(waterYear = as.character(waterYear)) %>%
  stns_daymonthyear<- df %>%
 mutate(Date = format(Date,"%d-%m-%Y")) %>%
  group_by(STATION_NUMBER) %>%
  fill(Value, .direction = 'downup') 
  return(stns_daymonthyear)
}

df_iha_ready <- lapply(df_list_rem_months, tidying_iha)

```
 
 
```{r}

#calculate IHA for full list
#currently calcs montly medians - can add monthly means by altering group 1 IHA function

scale_IHA <- function(data) {
df <- {{data}}
lst_stns <- split(df, df$STATION_NUMBER)
lst_stns_IHA <- lapply(lst_stns, calc_IHA)
IHA <- bind_rows(lst_stns_IHA, .id = "STATION_NUMBER")
}

 
 IHA_results_list <- lapply(df_iha_ready, function(df){
   function_result <- scale_IHA(df)
   function_result 
 })


 #ihaallresults <- bind_rows(IHA_results_list) #doesnt' give ecoregion name

#ecoreg <- lapply(names(resultslist), function(Ecoregion_name){
 # df<-lst[[Ecoregion_name]]
#  df$Ecoregion <- Ecoregion_name
 # df   })

#write.csv(ihaallresults, "iharesults.csv")

```


#B dates and ice variables  

```{r}


ymd <- function(data) {
  
  dfinput <- {{data}}
  results_ymd <- dfinput %>%
  mutate(Date = as.Date(Date, format = "%d-%m-%Y"))

}

df_ice_ready <- lapply(dflst, ymd)

#dflst
#df_iha_ready

```


```{r}

#Ice Variables group 1

scale_ice_g1 <- function(data) {
df <- {{data}}
lst_stns <- split(df, df$STATION_NUMBER)
lst_stns_IHA <- lapply(lst_stns, Group_1_ice_cover)
Ice_g1 <- bind_rows(lst_stns_IHA, .id = "STATION_NUMBER")
}

 
 Ice_g1_results_list <- lapply(df_ice_ready, function(df){
   function_result_ice_g1 <- scale_ice_g1(df)
   function_result_ice_g1 
 })

 
#ecoreg <- lapply(names(function_result_ice_g1), function(Ecoregion_name){
#  df<-lst[[Ecoregion_name]]
#  df$Ecoregion <- Ecoregion_name
 # df   })
#ihaallresults <- bind_rows(resultslist)

#write.csv(ihaallresults, "iharesults.csv")


```



```{r}

#Ice Variables group 2

scale_ice_g2 <- function(data) {
df <- {{data}}
lst_stns <- split(df, df$STATION_NUMBER)
lst_stns_IHA <- lapply(lst_stns, Group_2_freeze_thaw)
Ice_g1 <- bind_rows(lst_stns_IHA, .id = "STATION_NUMBER")
}


 Ice_g2_results_list <- lapply(df_ice_ready, function(df){
   function_result_ice_g2 <- scale_ice_g2(df)
   function_result_ice_g2
 })


#ecoreg <- lapply(names(function_result_ice_g2), function(Ecoregion_name){
#  df<-lst[[Ecoregion_name]]
#  df$Ecoregion <- Ecoregion_name
 # df   })
#ihaallresults <- bind_rows(resultslist)

#write.csv(ihaallresults, "iharesults.csv")

 
```
```{r}

STOP

```


#First and last B dates for each water year 
wy == 8

```{r}

#test dfs
tundra_10ND002_1980 <- tundra_10ND002 %>% filter(waterYear >= 1980)

wytest <- tundra_10ND002 %>% select(-c(waterYear))

tundra_10ND002_wytest <- add_water_year(wytest, 8)  %>%
filter(waterYear >= 1980)

```


```{r}

test <- tundra_10ND002_wytest %>%
  group_by(STATION_NUMBER, waterYear) %>%
  mutate(as.Date(Date)) %>%
  mutate(first_B = ifelse(Symbol == "B" & !is.na(Symbol), Date, NA),
         first_B = ifelse(any(!is.na(first_B)), min(first_B, na.rm = TRUE), NA),
         first_B = as.Date(first_B)) %>%
  mutate(last_B = ifelse(Symbol == "B" & !is.na(Symbol), Date, NA),
         last_B = ifelse(any(!is.na(last_B)), max(last_B, na.rm = TRUE), NA),
         last_B = as.Date(last_B)) 

#filter to get closed water data
test %>%
  group_by(STATION_NUMBER, waterYear) %>%
  filter(Date > first_B & Date < last_B) 
 
#filter to get open water data --> doesn't work
test %>%
  group_by(STATION_NUMBER, waterYear) %>%
  filter(Date > first_B & Date > last_B)


```


```{r}



```



```{r}



```


```{r}




```


#Next steps and questions:
```{r}

#annual variables:
#number missing dates, annual, sd etc.

#B dates: 
#mean, med, sd of flow data 1991-2021 (for each B date)
#number days, Bs/Es/As/NAs between first/last B date
#Ratio # days to Bs
#number missing flows before/after each B date
#number and type of flags before/after each B date
#mean, med, sd of first/last B date over 30 yrs and see if all stations are between this
#missing B dates 
#(rtn df with all of this)

#variables:
#OW: all IHA, annual vars
#CL: all IHA

#put in tableau?

#freshet:
#calculate freshet inflection - try shorter running means, first day when flow is 2x the previous day, see how long that happens for and do a midpoint
#clustering using regime shape?
#


```


```{r}



```


```{r}



```
 
 
 
```{r}



```

#Graphs: pre-removing months 

```{r, fig.height= 10, warning= FALSE}

#function to make list of dfs to plot
list_to_plot <- function(df){
  toplot <- df %>%
  dplyr::group_by(STATION_NUMBER) %>%
  dplyr::summarise(Year = unique(waterYear)) %>%
  arrange(STATION_NUMBER)
  return(toplot)
}

df_to_plot <- map(df_list, list_to_plot)

#make a plot object list of dfs 

plotlist <- list()
for (i in seq_along(df_to_plot)) {
  df_subset <- df_to_plot[[i]] #subset of df, by ecoregion

p <- ggplot(df_subset, aes(x = Year, y = STATION_NUMBER)) +
  geom_tile(fill = "blue", color = "white") +
  theme_classic() +
  ylab("") +
 ggtitle(paste("Data Availability - ", names(df_to_plot))[i]) +
    theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1, size= 8), plot.title = element_text(hjust = 0.5), axis.text.y = element_text(size = 5)) 
  plotlist[[i]] <- p     

}

names(plotlist) <- c('eastern_temp_forests', 'great_plains', 'hudson_plains', 'w_coast_forests', 'no_ecoregion', 'n_american_deserts', 'northern_forests', 'nw_forested_mtns', 'taiga', 'tundra')

```


```{r}

#save all plots to output file 
for (i in names(plotlist)){
  plot <- plotlist[[i]]
  filename <- paste0("output/", i, ".png")
  ggsave(filename, plot)
}

```

#Graphs: post-removing months 

```{r, fig.height= 10, warning= FALSE}

df_to_plot <- map(df_list_rem_months, list_to_plot)

#make a plot object list of dfs 

plotlist <- list()
for (i in seq_along(df_to_plot)) {
  df_subset <- df_to_plot[[i]] #subset of df, by ecoregion

p <- ggplot(df_subset, aes(x = Year, y = STATION_NUMBER)) +
  geom_tile(fill = "orange", color = "white") +
  theme_classic() +
  ylab("") +
 ggtitle(paste("Data Availability - ", names(df_to_plot))[i]) +
    theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1, size= 8), plot.title = element_text(hjust = 0.5), axis.text.y = element_text(size = 5)) 
  plotlist[[i]] <- p     

}

names(plotlist) <- c('eastern_temp_forests_2', 'great_plains_2', 'hudson_plains_2', 'w_coast_forests_2', 'no_ecoregion_2', 'n_american_deserts_2', 'northern_forests_2', 'nw_forested_mtns_2', 'taiga_2', 'tundra_2')


#save all plots to output file 
for (i in names(plotlist)){
  plot <- plotlist[[i]]
  filename <- paste0("output/", i, ".png")
  ggsave(filename, plot)
}

```



```{r}




```


```{r}

#Delete years with >3 missing days 

#delete missing data
#df_lst_del_yrs <-map(df_lst, calc_rle)

```

#random code 
```{r}

#next find all years available for each station
allstns <- df_list %>%
mutate(Date = ymd(Date),  Year = format_ISO8601(Date, precision = "y"))
filter(Year > 1969 & Year < 2024)

allstns_dt %>%
  group_by(STATION_NUMBER) %>%
  summarize(years_with_data = toString(unique(Year)))

#years with good data - delete years with >3 days a month of missing data, delete that whole year before doing calculations 


```

```{r}
#graph the data availability for all of the stations before deleting years with 3 days of data missing

graph_func <- function(data){
  ecoregion <- {{data}}[i]

yrs_full <- {{data}} %>%
  dplyr::group_by(STATION_NUMBER) %>%
  dplyr::summarise(Year = unique(waterYear))


#graphing data availability
p <- ggplot(yrs_full, aes(x = Year, y = STATION_NUMBER)) +
  geom_tile(fill = "lightgreen", color = "white") +
  theme_classic() +
  ylab("") +
  ggtitle("Data Availability - ") +
    theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1, size= 9), plot.title = element_text(hjust = 0.5))
return(p)
}


map(df_lst, graph_func) #cant title by the name of each df if do it this way, need to use a loop
```

```{r}
#make a bunch of dfs for plotting and then put them in a list 
et_test_full <- df_list$eastern_temp_forests %>%
  dplyr::group_by(STATION_NUMBER) %>%
  dplyr::summarise(Year = unique(waterYear)) %>%
  arrange(STATION_NUMBER)

c <- ceiling(n_distinct(unique(df_lst$Taiga$STATION_NUMBER))/30)

df_parts <- split(taiga_test_yrs, cut(seq_len(nrow(taiga_test_yrs)), breaks = c))

taiga1 <- df_parts[[1]]
taiga2 <- df_parts[[2]]
taiga3 <- df_parts[[3]]

df_list_toplot <- list(taiga1, taiga2, taiga3, tundra_test_yrs)
names(df_list_toplot) <- c('taiga1', 'taiga2', 'taiga3', 'tundra_test_yrs')


```


```{r}
#don't run this anymore, see function remove_months.... where just remove where there are more than three days of missing data
# function to filter out the seasonal months for seasonal stations only 
#filter_seas <- function(data) {
 # df <- {{data}} %>%
#  filter(Data.Quality == 'S' & month(Date) %in% c(3,4,5,6,7,8,9,10))
#}

#df_lst <- map(df_lst, filter_seas)
#df_lst_filter <- map(dflst, filter_seas)

```


```{r}

#sanity checks
n_distinct(stns$STATION_NUMBER) #1058
n_distinct(allstns$STATION_NUMBER) #1057

stns_ids <- stns %>% select(STATION_NUMBER) %>% distinct                                
allstns_ids <- allstns %>% select(STATION_NUMBER) %>% distinct

anti_join(stns_ids, allstns_ids)
# station 08LD001 not in master list of all stations, need to get this one or find out if need it or not 

```

```{r}


ggplot(diamonds, aes(carat, price)) + 
  geom_hex()
ggsave("output/diamonds.pdf")

write_csv(diamonds, "output/diamonds.csv")

summarise(STATION_NUMBER = paste(STATION_NUMBER,collapse = "','"))


```

