---
##RHBN STATIONS- TECHNICAL MEETING PREPARATION
---


```{r}
library(IHA)
library(tidyverse)
library(tidyhydat)
library(zoo)
library(lubridate)
library(ggplot2)
library(dataRetrieval)
```


```{r}
#delete later on
stn_07DA001 <- tidyhydat::hy_daily_flows(station_number = "07DA001") 
stn_07AA001 <- tidyhydat::hy_daily_flows(station_number = "07AA001") 

```

#NEED TO DO: convert to runoff

```{r}


```

#Data preparation 

1) Get a complete set of years for 1970-2020 (by year, not water year)

```{r}

tidying <- function(df){
  tidied <- df %>%
  group_by(STATION_NUMBER) %>%
  tidyr::complete(Date = seq.Date(as.Date("1969-08-01"), as.Date("2020-07-31"), by="day")) %>%
  mutate(Date = as.Date(Date)) %>%
  filter(Date >= "1969-08-01" & Date <= "2021-07-31") %>%
    ungroup()
  
  counts <- tidied %>%
   mutate(Year = year(Date)) %>%
  filter(Year > 1969 & Year < 2020) %>%
  group_by(STATION_NUMBER, Year) %>%
  summarize(count = n()) %>%
  filter(count < 365) 
  
 counts_yr <- tidied %>%
   mutate(Year = year(Date)) %>%
  filter(Year > 1969 & Year < 2020) %>%
  select(-c(Date, Value, Symbol, Parameter)) %>%
  distinct() %>%
  mutate(counts = n_distinct(Year)) %>%
  filter(counts < 50) 
  
  if (nrow(counts) > 0) {
    warning("There are STATION_NUMBER, Year groups that do not have 365 or 366 values:\n",
            paste0("STATION_NUMBER: ", counts$STATION_NUMBER, 
                       ", waterYear: ", counts$Year, 
                       ", Values: ", counts$count, collapse = "\n")) 
  }else {"all stations have 365 or 366 days per calendar Year"}
  
 
 if (nrow(counts_yr) > 0) {
    warning("There are STATION_NUMBER, Year groups that do not have 50 values:\n",
            paste0("STATION_NUMBER: ", counts$STATION_NUMBER, 
                       ", waterYear: ", counts$Year, 
                       ", Values: ", counts$count, collapse = "\n")) 
  } else {"all stations have 50 years"}
  
    rm(counts, counts_yr)

  return(tidied)

}


df_tidied <- tidying(stn_07AA001)

```


2) Get the water year
 and make sure there are 365-366 set of days for each WY and 50 water years  

```{r}


source("scripts/add_water_year.R")

month <- 10
df_wy <- add_water_year(df_tidied, month)


  counts <- df_wy %>%
  filter(waterYear > 1969 & waterYear < 2021) %>%
  group_by(STATION_NUMBER, waterYear) %>%
  summarize(count = n()) %>%
  filter(count < 365) 
  
  
 counts_yr <- df_wy %>%
   mutate(Year = year(Date)) %>%
  filter(Year > 1969 & Year < 2020) %>%
  select(-c(Date, Value, Symbol, Parameter)) %>%
  distinct() %>%
  mutate(counts = n_distinct(Year)) %>%
  filter(counts < 50) 
  
  if (nrow(counts) > 0) {
    warning("There are STATION_NUMBER, Year groups that do not have 365 or 366 values:\n",
            paste0("STATION_NUMBER: ", counts$STATION_NUMBER, 
                       ", waterYear: ", counts$waterYear, 
                       ", Values: ", counts$count, collapse = "\n")) 
  } else {"all stations have 365 or 366 days per chosen Water Year"}
 
 
 if (nrow(counts_yr) > 0) {
    warning("There are STATION_NUMBER, Year groups that do not have 50 values:\n",
            paste0("STATION_NUMBER: ", counts$STATION_NUMBER, 
                       ", waterYear: ", counts$Year, 
                       ", Values: ", counts$count, collapse = "\n")) 
  } else {"all stations have 50 years"}
  
    rm(counts, counts_yr)


```

3) Get the day of the year next

```{r}


calc_day_of_wyear <- function(data){
  
  grouped_df <- {{data}} %>%
    group_by(STATION_NUMBER, waterYear) %>%
    mutate(day_of_year = row_number()) %>%
    ungroup()
  
  counts <- grouped_df %>%
    group_by(STATION_NUMBER, waterYear) %>%
    summarize(count = n()) 
  
  wrong_num_groups <- counts %>% filter(count !=365 & count != 366)
  if(nrow(wrong_num_groups) > 0){
            warning("The following groups do not have 365 or 366 values:\n", 
            paste0("STATION_NUMBER: ", wrong_num_groups$STATION_NUMBER, 
                       ", waterYear: ", wrong_num_groups$waterYear, 
                       ", Values: ", wrong_num_groups$count, collapse = "\n"))
}  
  rm(counts, wrong_num_groups)
  return(grouped_df)

    }


df_doy <- calc_day_of_wyear(df_wy)


check_max <- df_doy %>%
  mutate(max = max(day_of_year)) %>%
  filter(max > 366) %>%
  distinct()

if(nrow(check_max) > 0){
            warning("There are days of the year that number more than 366, check the day of the year function")
}  else {"The days in the year are 365 or 366 days"}

rm(check_max) 

df_doy <- df_doy %>% filter(!waterYear %in% c(1969, 2021))

```

4) Get the weeks 1:52 for each year  

```{r}

calc_weeks <- function(data){
  
  df <- {{data}} %>%
    group_by(STATION_NUMBER, waterYear) %>%
  dplyr::mutate(weeks = rep(1:(ceiling(n()/7)), each = 7)[1:n()]) %>%  
  mutate(weeks = if_else(weeks == 53, 52, weeks)) %>% #make the one day week 53 be week 52. Issue is that then there are 9 days in week 52 in leap years. Could just omit feb 29th completely?
  ungroup()
}

df_ready <- calc_weeks(df_doy)

check_max <- df_ready %>%
  mutate(max = max(weeks)) %>%
  filter(max > 52) 

if(nrow(check_max) > 0){
            warning("There are more than 52 weeks in the year, check the week function or data")
} else {"all years have 52 weeks in a year"}


```

#label each year annual or seasonal
-what do we do if both conditions are false, ie if a year is not annual and not seasonal?

```{r}

#label each year as seasonal or annual

label_annual_seas <- function(x){
  df <- {{x}}
  df <- df %>% 
  group_by(STATION_NUMBER, waterYear) %>%
  mutate(allvals = all(!is.na(Value) & Value !=0),
  seasvals = all(!is.na(Value[day_of_year >= 90 & day_of_year <= 274]) & Value[day_of_year >= 90 & day_of_year <= 274] != 0),
         label = case_when(allvals ~ "annual", !allvals & seasvals ~ "seasonal", TRUE ~ NA_character_))
  
  df_check <- df %>% filter(allvals == FALSE & seasvals == FALSE & !is.na(label))
  
  if (nrow(df_check) > 0){
    warning("Strange, there are years that meet FALSE for both seasonal and annual definitions and yet have a seasonal or annual label. Maybe get rid of any years that have less than 365 days.")
  }else {"no issues"}
  rm(df_check)
  
  df_check_label <- df %>% filter(is.na(label) & !is.na(Value))
  if (nrow(df_check_label) >0) {
    warning("There are years that meet false for both seasonal and annual definitions. This data will be left out in some of the analyses that precede.")
  }
  #print(df_check_label)
  rm(df_check_label)
  return(df)
  
}

df_labeled <- label_annual_seas(df_ready)


```


Helper functions to check for if days missing in a month or year (no Value)

```{r}
#note, missing weeks defined by if there is a day missing for grouped station_number, year, week, not for weeks overall 

rem_missing_wks <- function(x, days){
  
  df <- {{x}}
  
  toremove <- df %>%
  group_by(STATION_NUMBER, waterYear, weeks) %>%
  select(-c(Date, Parameter, day_of_year, month)) %>%
  mutate(week_nas = sum(is.na(Value))) %>%
  filter(week_nas > {{days}}) %>%
  distinct()%>%
    ungroup()
  
  df_removed <- anti_join(df, toremove, by = c('STATION_NUMBER', 'waterYear', 'weeks'))

  return(df_removed)
}



rem_missing_months <- function(x, days){
  
  df <- {{x}} %>% mutate(month = month(Date))
  
  toremove <- df %>%
  group_by(STATION_NUMBER, waterYear, month) %>%
  select(-c(Date, Parameter, day_of_year, weeks)) %>%
  mutate(month_nas = sum(is.na(Value))) %>%
  filter(month_nas > {{days}}) %>%
  distinct() %>%
    ungroup()
  
  df_removed <- anti_join(df, toremove, by = c('STATION_NUMBER', 'waterYear', 'month'))

  return(df_removed)
}


#removes months with missing days and then years with missing months
#or do we instead want to remove years based on if there are < 365 days ?

rem_missing_years <- function(x, days){
  df <- {{x}}
  df_mnths_removed <- rem_missing_months(df, {{days}})
  
  toremove <- df_mnths_removed %>%
  group_by(STATION_NUMBER, waterYear) %>%
  summarize(month = n_distinct(month)) %>%
  filter(month < 12) %>%
    ungroup()
  
  df_removed <- anti_join(df_mnths_removed, toremove, by = c('STATION_NUMBER', 'waterYear'))

  return(df_removed)
  
}

```


```{r}




```


#Annual variables 

Annual and monthly mean & median

```{r}

#note: only do if label == 'annual'

calc_mn_md_annual <- function(data){
  input_data <- {{data}}
  results <- input_data %>%
    dplyr::group_by(STATION_NUMBER, waterYear) %>%
    dplyr::summarize(annual_mean = round(mean(Value), 1), annual_median = round(median(Value),1), annual_sd = sd(Value)) %>% ungroup()
  return(results)
}


calc_mn_md_month <- function(data){
  input_data <- {{data}}
  results <- input_data %>%
      mutate(month = lubridate::month(Date, label = TRUE, abbr = TRUE)) %>%
    dplyr::group_by(STATION_NUMBER, waterYear, month) %>%
    dplyr::summarize(monthly_mean = round(mean(Value),1), monthly_median = round(median(Value), 1), monthly_std = round(sd(Value), 1)) %>%
  pivot_wider(names_from = month, values_from = c(monthly_mean, monthly_median, monthly_std), names_prefix = "") %>% ungroup()
    return(results)
}

yrs_all <- 1970:2020

df_yrly_means_med  <- df_labeled %>% filter(label == 'annual') %>% calc_mn_md_annual() %>%
  group_by(STATION_NUMBER) %>%  complete(waterYear = yrs_all)  %>% ungroup()

df_mnthly_means_med <- df_labeled %>% filter(label == 'annual') %>% rem_missing_years(days = 3)  %>% calc_mn_md_month()%>%
  group_by(STATION_NUMBER) %>%  complete(waterYear = yrs_all)  %>% ungroup()

```


Weekly mean & median


```{r}

calc_mn_md_week <- function(x){
  df <- {{x}}
  results <- df %>%
    group_by(STATION_NUMBER, weeks) %>%
   summarize(weekly_mean = round(mean(Value),1), weekly_median = round(median(Value), 1))
    return(results)
}

#issue is if removing weeks, how does this effect the final outcome, and do we need to know how many weeks were removed? Does there need to be a minimum amount of weeks available to get the means/medians?

df_labeled %>% rem_missing_wks(days = 1) %>% calc_mn_md_week()

```


IHA variables - annual or seasonal, as needed 
```{r}

stn_cln <- df_labeled %>%
  mutate(Date = as.Date(Date)) %>%
  filter(waterYear > 1970) %>%
  filter(waterYear < 2021) %>%
  filter(label == 'annual') #change 

stns_daymonthyear <- stns_ready %>%
  mutate(Date = format(Date,"%d-%m-%Y")) %>%
  group_by(STATION_NUMBER) 

lst_stns <- split(stns_daymonthyear, stns_daymonthyear$STATION_NUMBER )

lst_stns_IHA <- lapply(lst_stns, calc_IHA)

IHA <- bind_rows(lst_stns_IHA, .id = "STATION_NUMBER")

View(IHA)

```


#trying to get the Median duration of Min pulses within each year
(#min pulses)
-date of 90 day magnitude 


```{r}

stn_cln <- df_labeled %>%
  mutate(Date = as.Date(Date)) %>%
  filter(waterYear > 1970) %>%
  filter(waterYear < 2021) %>%
  filter(label == 'annual')  

stns_daymonthyear <- stn_cln %>%
  mutate(Date = format(Date,"%d-%m-%Y"))  %>%
      dplyr::select(Date, Value)

flow_data <- zoo(stns_daymonthyear$Value, order.by = as.Date(as.character(stns_daymonthyear$Date), format = "%d-%m-%Y"))

lst_stns <- split(stns_daymonthyear, stns_daymonthyear$STATION_NUMBER )

lst_stns_IHA <- lapply(lst_stns,group2 )

IHA <- bind_rows(lst_stns_IHA, .id = "STATION_NUMBER")


group2(flow_data, year = "water", mimic.tnc = TRUE)

group2_output <- group2(flow_data, year = "water", mimic.tnc = TRUE)
group2_output <- group2_output[,-1]

group2test(flow_data)

```

Annual- first b date, last b date, magnitude
- # days open water, number days B date

```{r}


```

Annual - ice effected vars
```{r}


```

Annual - open water vars 
```{r}


```



#extra code- remove months with > 3 missing days 
```{r}

monthly_missing_days <- df_labeled %>%
  group_by(STATION_NUMBER, waterYear, month) %>%
  summarise(missing_days = sum(is.na(Value))) %>%
  ungroup()

# Identify months with more than 3 missing days
months_to_flag <- monthly_missing_days %>%
  filter(missing_days > 3) %>%
  select(STATION_NUMBER, waterYear, month)

# Create flagged dataset
df_flagged <- df_labeled %>%
  semi_join(months_to_flag, by = c("STATION_NUMBER", "waterYear", "month"))

# Update the original dataframe
df_updated <- df_labeled %>%
  anti_join(df_flagged, by = c("STATION_NUMBER", "waterYear", "month")) %>%
  bind_rows(df_flagged %>%
              mutate(Value = NA))

# Check the updated dataframe
print(df_updated)

```

```{r}
```

```{r}
```

```{r}
```

```{r}
```


```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```


```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```


```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```
